Fast and exact visibility on digitized shapes and application to saliency-aware normal estimation

### 2025-04-15 DGMM2025/paper.tex
graphs and visibility test 2d example

\usepackage{changes}
\usepackage{algorithm}
\usepackage[noend]{algpseudocode}
\usepackage{varwidth}
\usepackage{mathtools}
\usepackage{pgfplots}
\setauthormarkup{}
\definechangesauthor[name=jo, color=magenta]{JO}
\definechangesauthor[name=rom, color=teal]{ROM}

@@

        \usetikzlibrary {arrows.meta,decorations.shapes}

@@

@@

    \begin{algorithm}
        \caption{Given 2 lists of intervals $K$ and $L$, find $K \cap L$, the intersection of those 2 lists}
        \label{alg:intersection}
        \begin{algorithmic}
            \Function{Intersection ($\cap$)}{\text{K, L}: Intervals}
                \State $R \gets \emptyset$
                \State $k, l \gets 0$
                \While{$k < K.nbIntervals \And l < L.nbIntervals$}
                    \State $[a,b] \gets K[k]$
                    \State $[c,d] \gets L[l]$
                    \State $e \gets \max(a, c)$
                    \State $f \gets \min(b, d)$
                    \If{$e \leq f$}\State $R.append([e, f])$
                    \EndIf
                    \If{$b \leq d$}\State $k \gets k+1$
                    \EndIf
                    \If{$d \leq b$}\State $l \gets l+1$
                    \EndIf
                \EndWhile
                \State \Return $R$
            \EndFunction
        \end{algorithmic}
    \end{algorithm}

@@

    \pgfplotsset{width=13cm}
    \begin{tikzpicture}
        \centering
        \begin{axis}[
            xlabel={Gridstep},
            x dir=reverse,
            ylabel={Mean distance},
            legend pos=north west,
            ymajorgrids=true,
            grid style=dashed,
            ymode=log,
            log ticks with fixed point,
        ]

            \addplot[
                color=cyan,
                mark=square,
            ]
            coordinates {
                (1,5.76973)(0.875,6.03836)(0.75,6.22426)(0.625,6.39012)(0.5,6.51014)(0.375,6.63571)(0.25, 6.68132)
            };
            \addlegendentry{sphere9 radius 10}
            \addplot[
                color=red,
                mark=square,
            ]
            coordinates {
                (1,5.91022)(0.875,6.10641)(0.75,6.14175)(0.625,6.20202)(0.5,6.28637)(0.375,6.44418)(0.25, 6.56938)
            };
            \addlegendentry{goursat radius 10}
            \addplot[
                color=red,
                mark=triangle,
            ]
            coordinates {
                (1,9.04228)(0.875,10.1319)(0.75,10.2837)(0.625,10.6125)(0.5,10.8969)(0.375,11.3186)(0.25, 11.8382)
            };
            \addlegendentry{radius 20}
            \addplot[
                color=red,
                mark=halfcircle,
            ]
            coordinates {
                (1,9.22083)(0.875,10.6476)(0.75,11.5976)(0.625,12.7428)(0.5,13.9534)(0.375,14.73)(0.25,15.8425)
            };
            \addlegendentry{radius 30}
            \addplot[
                color=violet,
                mark=square,
            ]
            coordinates {
                (1,5.36873)(0.875,5.43148)(0.75,5.47206)(0.625,5.67227)(0.5,5.81554)(0.375,6.03961)(0.25,6.31671)
            };
            \addlegendentry{torus}
            \addplot[
                color=violet,
                mark=triangle,
            ]
            coordinates {
                (1,6.42158)(0.875,6.36929)(0.75,6.94675)(0.625,7.65529)(0.5,8.65652)(0.375,9.05402)(0.25,10.0784)
            };
            \addplot[
                color=violet,
                mark=halfcircle,
            ]
            coordinates {
                (1,6.44119)(0.875,6.43563)(0.75,7.00852)(0.625,7.71613)(0.5,9.1544)(0.375,9.84299)(0.25,12.241)
            };
        \end{axis}
    \end{tikzpicture}

>>>>

\usepackage{changes}
\usepackage{algorithm}
\usepackage[noend]{algpseudocode}
\usepackage{varwidth}
\usepackage{mathtools}
\usepackage{pgfplots}
\usetikzlibrary {arrows.meta,decorations.shapes}
\setauthormarkup{}
\definechangesauthor[name=jo, color=magenta]{JO}
\definechangesauthor[name=rom, color=teal]{ROM}

@@

        % \usetikzlibrary {arrows.meta,decorations.shapes}

@@

    \begin{figure}
        \centering
        \begin{tabular}{c c}
            \begin{tikzpicture}[scale=0.75]
                \draw[step=0.5,lightgray,thin,xshift=-1cm,yshift=-1cm] (-0.75,-0.25) grid (12.25,3.25);
                \draw[-{Stealth[length=3mm]}=black] (-1.5,-1) -- (-1.5,2.5);
                \draw decorate [decoration={crosses,transform={rotate=45},shape size=1.5mm,segment length=21.5pt}] {(-1.5,-1) -- (-1.5,2)};
                \filldraw[black] (0,0) circle (0.05);
                \filldraw[black] (1,0) circle (0.05);
                \filldraw[black] (2,0) circle (0.05);
                \filldraw[black] (2,1) circle (0.05);
                \filldraw[black] (3,1) circle (0.05);
                \filldraw[black] (4,1) circle (0.05);
                \filldraw[black] (5,1) circle (0.05);
                \filldraw[black] (5,0) circle (0.05);
                \filldraw[black] (6,0) circle (0.05);
                \filldraw[black] (7,0) circle (0.05);
                \filldraw[black] (7,1) circle (0.05);
                \filldraw[black] (8,1) circle (0.05);
                \filldraw[black] (8,0) circle (0.05);
                \filldraw[black] (9,0) circle (0.05);
                \filldraw[black] (10,0) circle (0.05);
                \filldraw[black] (10,1) circle (0.05);
                \draw[thick,gray, arrows = {Bracket[sharp]-Bracket[sharp]}] (-0.6,-0.5)--(2.6,-0.5);
                \draw[thick,gray, arrows = {Bracket[sharp]-Bracket[sharp]}] (4.4,-0.5)--(10.6,-0.5);
                \draw[thick,gray, arrows = {Bracket[sharp]-Bracket[sharp]}] (-0.6,0)--(2.6,0);
                \draw[thick,gray, arrows = {Bracket[sharp]-Bracket[sharp]}] (4.4,0)--(10.6,0);
                \draw[thick,gray, arrows = {Bracket[sharp]-Bracket[sharp]}] (-0.6,0.5)--(10.6,0.5);
                \draw[thick,gray, arrows = {Bracket[sharp]-Bracket[sharp]}] (1.4,1)--(5.6,1);
                \draw[thick,gray, arrows = {Bracket[sharp]-Bracket[sharp]}] (6.4,1)--(8.6,1);
                \draw[thick,gray, arrows = {Bracket[sharp]-Bracket[sharp]}] (9.4,1)--(10.6,1);
                \draw[thick,gray, arrows = {Bracket[sharp]-Bracket[sharp]}] (1.4,1.5)--(5.6,1.5);
                \draw[thick,gray, arrows = {Bracket[sharp]-Bracket[sharp]}] (6.4,1.5)--(8.6,1.5);
                \draw[thick,gray, arrows = {Bracket[sharp]-Bracket[sharp]}] (9.4,1.5)--(10.6,1.5);
                \draw[thick, black] (0,0) -- (2,0) -- (2,1) -- (5,1) -- (5,0) -- (7,0) -- (7,1) -- (8,1) -- (8,0) -- (10,0) -- (10,1);
                \draw[thick,dashed] (6,1) -- (6,2);
                \draw[thick,dashed] (9,1) -- (9,2);
                \draw[thick,dashed] (-1,-1) -- (-1,1) -- (1,1) -- (1,2) -- (11,2) -- (11,-1) -- (4,-1) -- (4,0) -- (3,0) -- (3,-1) -- (-1,-1);

            \end{tikzpicture} &
            \begin{tikzpicture}[scale=0.75]
                \draw[step=0.5,lightgray,thin,xshift=-1cm,yshift=-1cm] (-0.75,-0.25) grid (4.25,3.25);
                \draw[-{Stealth[length=3mm]}=black] (-1.5,-1) -- (-1.5,2.5);
                \draw decorate [decoration={crosses,transform={rotate=45},shape size=1.5mm,segment length=21.5pt}] {(-1.5,-1) -- (-1.5,2)};
                \filldraw[black] (0,0) circle (0.05);
                \filldraw[black] (2,1) circle (0.05);
                \draw[thick, black] (0,0) -- (2,1);
                \draw[thick,dashed] (-1,-1) -- (-1,1) -- (1,1) -- (1,2) -- (3,2) -- (3,0) -- (1,0) --  (1,-1) -- (-1,-1);
                \draw[thick,gray, arrows = {Bracket[sharp]-Bracket[sharp]}] (-0.6,-0.5)--(0.6,-0.5);
                \draw[thick,gray, arrows = {Bracket[sharp]-Bracket[sharp]}] (-0.6,0)--(0.6,0);
                \draw[thick,gray, arrows = {Bracket[sharp]-Bracket[sharp]}] (-0.6,0.5)--(2.6,0.5);
                \draw[thick,gray, arrows = {Bracket[sharp]-Bracket[sharp]}] (1.4,1)--(2.6,1);
                \draw[thick,gray, arrows = {Bracket[sharp]-Bracket[sharp]}] (1.4,1.5)--(2.6,1.5);
            \end{tikzpicture}\\
            \hline
            \multicolumn{2}{c}{
                \begin{tikzpicture}[scale=0.75]
                    \foreach \y in {0,0.5,...,10.5} {
                        \draw[thin, lightgray] (-1.25,\y) -- (11.25,\y);
                    }
                    \foreach \y in {0,2,...,10} {
                        \draw[blue] (-1.25,\y) -- (11.25,\y);
                    }

                    \draw[thick,green, arrows = {Bracket[sharp]-Bracket[sharp]}] (-0.6,9.5)--(0.6,9.5);
                    \draw[thick,arrows = {Bracket[sharp]-Bracket[sharp]}] (-0.6,9)--(2.6,9);
                    \draw[thick,arrows = {Bracket[sharp]-Bracket[sharp]}] (4.4,9)--(10.6,9);
                    \draw[thick,red, arrows = {Bracket[sharp]-Bracket[sharp]}] (-0.6,8.5)--(1.6,8.5);
                    \draw[thick,red, arrows = {Bracket[sharp]-Bracket[sharp]}] (4.4,8.5)--(9.6,8.5);

                    \draw[thick,green, arrows = {Bracket[sharp]-Bracket[sharp]}] (-0.6,7.5)--(0.6,7.5);
                    \draw[thick,arrows = {Bracket[sharp]-Bracket[sharp]}] (-0.6,7)--(2.6,7);
                    \draw[thick,arrows = {Bracket[sharp]-Bracket[sharp]}] (4.4,7)--(10.6,7);
                    \draw[thick,red, arrows = {Bracket[sharp]-Bracket[sharp]}] (-0.6,6.5)--(1.6,6.5);
                    \draw[thick,red, arrows = {Bracket[sharp]-Bracket[sharp]}] (4.4,6.5)--(9.6,6.5);

                    \draw[thick,green, arrows = {Bracket[sharp]-Bracket[sharp]}] (-0.6,5.5)--(2.6,5.5);
                    \draw[thick,arrows = {Bracket[sharp]-Bracket[sharp]}] (-0.6,5)--(10.6,5);
                    \draw[thick,red, arrows = {Bracket[sharp]-Bracket[sharp]}] (-0.6,4.5)--(1.6,4.5);
                    \draw[thick,red, arrows = {Bracket[sharp]-Bracket[sharp]}] (4.4,4.5)--(7.6,4.5);

                    \draw[thick,green, arrows = {Bracket[sharp]-Bracket[sharp]}] (1.4,3.5)--(2.6,3.5);
                    \draw[thick, arrows = {Bracket[sharp]-Bracket[sharp]}] (1.4,3)--(5.6,3);
                    \draw[thick, arrows = {Bracket[sharp]-Bracket[sharp]}] (6.4,3)--(8.6,3);
                    \draw[thick, arrows = {Bracket[sharp]-Bracket[sharp]}] (9.4,3)--(10.6,3);
                    \draw[thick,red, arrows = {Bracket[sharp]-Bracket[sharp]}] (-0.6,2.5)--(1.6,2.5);
                    \draw[thick,red, arrows = {Bracket[sharp]-Bracket[sharp]}] (4.4,2.5)--(5.6,2.5);
                    \draw[thick,red, arrows = {Bracket[sharp]-Bracket[sharp]}] (7.4,2.5)--(7.6,2.5);

                    \draw[thick,green, arrows = {Bracket[sharp]-Bracket[sharp]}] (1.4,1.5)--(2.6,1.5);
                    \draw[thick, arrows = {Bracket[sharp]-Bracket[sharp]}] (1.4,1)--(5.6,1);
                    \draw[thick, arrows = {Bracket[sharp]-Bracket[sharp]}] (6.4,1)--(8.6,1);
                    \draw[thick, arrows = {Bracket[sharp]-Bracket[sharp]}] (9.4,1)--(10.6,1);
                    \draw[thick,red, arrows = {Bracket[sharp]-Bracket[sharp]}] (-0.6,0.5)--(1.6,0.5);
                    \draw[thick,red, arrows = {Bracket[sharp]-Bracket[sharp]}] (4.4,0.5)--(5.6,0.5);
                    \draw[thick,red, arrows = {Bracket[sharp]-Bracket[sharp]}] (7.4,0.5)--(7.6,0.5);

                \end{tikzpicture}}
        \end{tabular}
        \caption{Evolution of the visibility check algorithm for a 2,1 vector. Green are the vector lattices, black are the figure lattices, red are the current intervals of positions where the visibility is still possible. The last red intervals are the visible positions. We travel the lattices from bottom-up. }
        \label{fig:enter-label}
    \end{figure}

@@

    \begin{algorithm}
        \caption{Given 2 lists of intervals $K$ and $L$, find $K \cap L$, the intersection of those 2 lists}
        \label{alg:intersection}
        \begin{algorithmic}
            \Function{Intersection ($\cap$)}{\text{K, L}: Intervals}
                \State $R \gets \emptyset$
                \State $k, l \gets 0$
                \While{$k < K.nbIntervals \And l < L.nbIntervals$}
                    \State $[a,b] \gets K[k]$
                    \State $[c,d] \gets L[l]$
                    \State $e \gets \max(a, c)$
                    \State $f \gets \min(b, d)$
                    \If{$e \leq f$}
                        \State $R.append([e, f])$
                    \EndIf
                    \If{$b \leq d$}
                        \State $k \gets k+1$
                    \EndIf
                    \If{$d \leq b$}
                        \State $l \gets l+1$
                    \EndIf
                \EndWhile
                \State \Return $R$
            \EndFunction
        \end{algorithmic}
    \end{algorithm}

@@

    \pgfplotsset{width=13cm}
    \begin{figure}
        \begin{tikzpicture}
            \centering
            \begin{axis}[
                xlabel={Gridstep},
                x dir=reverse,
                ylabel={Mean distance},
                legend pos=north west,
                ymajorgrids=true,
                grid style=dashed,
                ymode=log,
                log ticks with fixed point,
            ]

                \addplot[
                    color=cyan,
                    mark=square,
                ]
                coordinates {
                    (1,5.76973)(0.875,6.03836)(0.75,6.22426)(0.625,6.39012)(0.5,6.51014)(0.375,6.63571)(0.25,6.68132)
                };
                \addlegendentry{sphere9 radius 10}
                \addplot[
                    color=red,
                    mark=square,
                ]
                coordinates {
                    (1,5.91022)(0.875,6.10641)(0.75,6.14175)(0.625,6.20202)(0.5,6.28637)(0.375,6.44418)(0.25,6.56938)
                };
                \addlegendentry{goursat radius 10}
                \addplot[
                    color=red,
                    mark=triangle,
                ]
                coordinates {
                    (1,9.04228)(0.875,10.1319)(0.75,10.2837)(0.625,10.6125)(0.5,10.8969)(0.375,11.3186)(0.25,11.8382)
                };
                \addlegendentry{radius 20}
                \addplot[
                    color=red,
                    mark=halfcircle,
                ]
                coordinates {
                    (1,9.22083)(0.875,10.6476)(0.75,11.5976)(0.625,12.7428)(0.5,13.9534)(0.375,14.73)(0.25,15.8425)
                };
                \addlegendentry{radius 30}
                \addplot[
                    color=violet,
                    mark=square,
                ]
                coordinates {
                    (1,5.36873)(0.875,5.43148)(0.75,5.47206)(0.625,5.67227)(0.5,5.81554)(0.375,6.03961)(0.25,6.31671)
                };
                \addlegendentry{torus}
                \addplot[
                    color=violet,
                    mark=triangle,
                ]
                coordinates {
                    (1,6.42158)(0.875,6.36929)(0.75,6.94675)(0.625,7.65529)(0.5,8.65652)(0.375,9.05402)(0.25,10.0784)
                };
                \addplot[
                    color=violet,
                    mark=halfcircle,
                ]
                coordinates {
                    (1,6.44119)(0.875,6.43563)(0.75,7.00852)(0.625,7.71613)(0.5,9.1544)(0.375,9.84299)(0.25,12.241)
                };
            \end{axis}
        \end{tikzpicture}
        \caption{Mean Visibility Distance as a Function of Grid Resolution}
    \end{figure}
    \begin{figure}
        \begin{tikzpicture}
            \centering
            \begin{axis}[
                xlabel={Gridstep},
                x dir=reverse,
                ylabel={Duration (s)},
                legend pos=north west,
                ymajorgrids=true,
                grid style=dashed,
                ymode=log,
                log ticks with fixed point,
            ]

                \addplot[
                    color=cyan,
                    mark=square,
                ]
                coordinates {
                    (1,0.513)(0.875,0.585)(0.75,0.897)(0.625,1.401)(0.5,2.441)(0.375,4.664)(0.25,11.420)
                };
                \addlegendentry{sphere9 radius 10}
                \addplot[
                    color=red,
                    mark=square,
                ]
                coordinates {
                    (1,0.440)(0.875,0.665)(0.75,0.823)(0.625,1.428)(0.5,2.463)(0.375,4.697)(0.25,11.766)
                };
                \addlegendentry{goursat radius 10}
                \addplot[
                    color=red,
                    mark=triangle,
                ]
                coordinates {
                    (1,1.389)(0.875,2.567)(0.75,3.661)(0.625,7.123)(0.5,13.881)(0.375,28.913)(0.25,77.360)
                };
                \addlegendentry{radius 20}
                \addplot[
                    color=red,
                    mark=halfcircle,
                ]
                coordinates {
                    (1,2.172)(0.875,3.747)(0.75,5.565)(0.625,13.180)(0.5,33.088)(0.375,80.584)(0.25,240.546)
                };
                \addlegendentry{radius 30}
                \addplot[
                    color=violet,
                    mark=square,
                ]
                coordinates {
                    (1,0.057)(0.875,0.081)(0.75,0.097)(0.625,0.197)(0.5,0.330)(0.375,0.800)(0.25,2.253)
                };
                \addlegendentry{torus}
                \addplot[
                    color=violet,
                    mark=triangle,
                ]
                coordinates {
                    (1,0.248)(0.875,0.294)(0.75,0.339)(0.625,0.620)(0.5,1.110)(0.375,2.643)(0.25,9.369)
                };
                \addplot[
                    color=violet,
                    mark=halfcircle,
                ]
                coordinates {
                    (1,0.837)(0.875,0.880)(0.75,0.942)(0.625,1.316)(0.5,2.142)(0.375,5.046)(0.25,20.418)
                };
            \end{axis}
        \end{tikzpicture}
        \caption{Computation Time of Visibility as a Function of Grid Resolution}
    \end{figure}
